$5e^{x+y}$: A Math Aware Search Engine (for CDS)

### 2014-11-04 docs/thesis/chap-clustering.tex
First load of the JCDL paper files.

\small{{\codefont ./mahout canpoy -t1 0.15 -t2 0.1 -i ../results/vectors -o ../results/clustering/canopy -dm org.apache.mahout.common.distance.CosineDistanceMeasure}}

>>>>

\small{{\codefont ./mahout canopy -t1 0.15 -t2 0.1 -i ../results/vectors -o ../results/clustering/canopy -dm org.apache.mahout.common.distance.CosineDistanceMeasure}}

### 2014-11-04 jcdl2015/paper/section-evaluation.tex
First load of the JCDL paper files.

\chapter{Evaluation}
When building an IR system there are two main aspects that should be considered. One is the quality of the presented results and the second one is the performance of the system since web users currently tolerate some milliseconds of latency for retrieving results.

\section{Dataset}
Our dataset consists on 12234 latex documents fetched from ArXiV. This documents correspond to the latest records harvested for CDS. The documents were processed using LateXML as described in the extraction process. From the given dataset, the simpler equations were filtered out (The ones containing one or two elements). The total number of indexed equations is 1412297.

\section{Setup}

@@

\begin{longtable}{|m{4cm}|>
{\centering\arraybackslash}m{4cm}|>
{\centering\arraybackslash}m{4cm}|
}
\hline 
\textbf{Software} & 
\textbf{Version} 
\\
\hline
Operative System & Linux Mint 16 Petra (Ubuntu based) \\ \hline
Linux Kernel & 3.11.0-15 \\ \hline
Java Virtual Machine & 1.7.0-45 \\ \hline
Apache Solr & 4.6 \\ \hline
Python & 2.7.5+ \\ \hline
\caption{Software used during evaluations}
\label{sw_used}
\end{longtable}

@@

\begin{longtable}{|c|c|c|c|}
\hline 
\textbf{Scenario} & \textbf{Top 5} & \textbf{Top 10} & \textbf{Top 20} \\ 
\hline 
N & 0.86 & 0.79 & 0.78 \\ 
\hline 
N+S & 0.9 & 0.85 & 0.77 \\ 
\hline 
N+S+NN & 0.92 & 0.94 & 0.88 \\ 
\hline
N+S+NN (BM25) & 0.92 & 0.94 & 0.87 \\ 
\hline

\caption{Loose Precision}
\label{loose_precision}
\end{longtable}

\begin{longtable}{|c|c|c|c|}
\hline 
\textbf{Scenario} & \textbf{Top 5} & \textbf{Top 10} & \textbf{Top 20} \\ 
\hline 
N & 0.76 & 0.61 & 0.50 \\ 
\hline 
N+S & 0.84 & 0.70 & 0.56 \\ 
\hline 
N+S+NN & 0.84 & 0.83 & 0.65 \\ 
\hline 
N+S+NN (BM25) & 0.84 & 0.83 & 0.67 \\ 
\hline

\caption{Strict Precision}
\label{stric_precision}
\end{longtable}

\begin{longtable}{|c|c|c|c|}
\hline 
\textbf{Scenario} & \textbf{Top 5} & \textbf{Top 10} & \textbf{Top 20} \\ 
\hline 
N & 7.17 & 9.81 & 12.00 \\ 
\hline 
N+S & 7.61 & 10.7 & 14.59 \\ 
\hline 
N+S+NN & 7.66 & 11.81 & 16.32 \\ 
\hline 
N+S+NN (BM25) & 7.66 & 11.82 & 16.42 \\ 
\hline
\caption{Discounted Cumulative Gain}
\label{dcg}
\end{longtable}

@@

\section{Indexing Performance}
For this set of evaluations, we were only interested in observing how the performance of the indexing stage is affected by the different types of features that are employed. We selected the first 100 records from our dataset and measured the total time it required to index all the elements in Solr using a python script and the solrpy library to do the communication. The default similarity in Solr was used. We also measured the index size at the end of each experiment.
Our four scenarios were organized as follows:
We tested the same three scenarios as for the quality results and we added a fourth one N+S+FN where we used the fully normalized form of the expressions as opposed to the standard simplified form in the N+S+NN scenario.

@@

\begin{longtable}{|c|c|>
{\centering\arraybackslash}m{2cm}|>
{\centering\arraybackslash}m{2cm}|>
{\centering\arraybackslash}m{2cm}|}
\hline 
\textbf{Feature Set} & \textbf{Total time (sec)} & \textbf{Index Size (kb)} & \textbf{Avg. time per equation} & \textbf{Avg. size of equation (kb)} \\ 
\hline 
N & 58 & 13926 & 0.00524 & 1.237 \\ 
\hline 
N+S & 884 & 14438 & 0.0785 & 1.282 \\ 
\hline 
N+S+NN & 1273 & 25780 & 0.113 & 2.345 \\ 
\hline 
N+S+FN & 1377 & 25548 & 0.122 & 2.269 \\ 
\hline
\caption{Indexing performance}
\label{indexing_performance}
\end{longtable}

\section{Querying Performance}
Similarly, we were also interested in evaluating how the query performance of the queries was affected with the different scenarios. For this set of queries we used the complete dataset, and evaluated the N and the N+S+NN scenarios for queries of different length.

\begin{longtable}{|>{\centering\arraybackslash}m{5cm}|>
{\centering\arraybackslash}m{4cm}|>
{\centering\arraybackslash}m{4cm}|
}
\hline 
\textbf{Expression Length(Number of characters in MathML)}& \textbf{N} & \textbf{N+S+NN} \\ \hline 
70 & 0.178 & 0.327  \\ \hline
198 & 0.179 & 0.810  \\ \hline
452 & 0.323 & 1.692  \\ \hline
771 & 0.393 & 2.066  \\ \hline

\caption{Query performance in seconds}
\label{query_performance}
\end{longtable}

@@

\section{Alternative Metrics}
According to some experimental evaluations presented in \cite{tree_comparison}, Tree Edit Distance provides an interesting alternative metric to the standard vector space model using angular distance. In this work, authors present an improvement in the quality of retrieved results compared to the classic approach of indexing the single leaves of the tree and a subset of the possible sub-trees. For evaluating the feasibility of this approach we compared a set of equations using both metrics: Tree edit distance (Using Zheng and Shasha's algorithm implemented in python \cite{tree_distance_python}.) and a standard angular distance that was not heavily optimized. We compared elapsed times for expressions of different length.

\begin{longtable}{|>{\centering\arraybackslash}m{5cm}|>
{\centering\arraybackslash}m{4cm}|>
{\centering\arraybackslash}m{4cm}|
}
\hline 
\textbf{Expression Length (Number of leaf nodes)}& \textbf{Angular Distance Time (msec)} & \textbf{Tree Edit Distance Time(msec)} \\ \hline 
3 & 0.01195 & 2.65  \\ \hline
8 & 0.0142 & 9.97  \\ \hline
7 & 0.0129 & 9.165  \\ \hline
12 & 0.0194 & 15.15  \\ \hline
18 & 0.0201 & 29.5  \\ \hline
20 & 0.0245 & 27.75  \\ \hline
22 & 0.0264 & 32.4  \\ \hline
24 & 0.02502 & 32.1 \\ \hline

\caption{Distance computation using two different metrics}
\label{comparison_tree_times}
\end{longtable}

>>>>

\section{Evaluation}
When building an IR system there are two main aspects that should be considered. One is the quality of the presented results and the second one is the performance of the system since web users currently tolerate some milliseconds of latency for retrieving results.

\subsection{Dataset}
Our dataset consists on 12234 latex documents fetched from ArXiV. This documents correspond to the latest records harvested for CDS. The documents were processed using LateXML as described in the extraction process. From the given dataset, the simpler equations were filtered out (The ones containing one or two elements). The total number of indexed equations is 1412297.

\subsection{Setup}

@@

@@

@@

\subsection{Indexing Performance}
For this set of evaluations, we were only interested in observing how the performance of the indexing stage is affected by the different types of features that are employed. We selected the first 100 records from our dataset and measured the total time it required to index all the elements in Solr using a python script and the solrpy library to do the communication. The default similarity in Solr was used. We also measured the index size at the end of each experiment.
Our four scenarios were organized as follows:
We tested the same three scenarios as for the quality results and we added a fourth one N+S+FN where we used the fully normalized form of the expressions as opposed to the standard simplified form in the N+S+NN scenario.

@@

\subsection{Querying Performance}
Similarly, we were also interested in evaluating how the query performance of the queries was affected with the different scenarios. For this set of queries we used the complete dataset, and evaluated the N and the N+S+NN scenarios for queries of different length.

@@

\subsection{Alternative Metrics}
According to some experimental evaluations presented in \cite{tree_comparison}, Tree Edit Distance provides an interesting alternative metric to the standard vector space model using angular distance. In this work, authors present an improvement in the quality of retrieved results compared to the classic approach of indexing the single leaves of the tree and a subset of the possible sub-trees. For evaluating the feasibility of this approach we compared a set of equations using both metrics: Tree edit distance (Using Zheng and Shasha's algorithm implemented in python \cite{tree_distance_python}.) and a standard angular distance that was not heavily optimized. We compared elapsed times for expressions of different length.

### 2014-11-04 jcdl2015/paper/section-related_work.tex
First load of the JCDL paper files.

\chapter{Related Work}
\label{chap-related_work}
\section{Status of Mathematical Information Retrieval}
Even though Information Retrieval has been a research field from around the 1950s where the first metrics for such systems were developed and in the 1960s where SMART (System for Mechanical Analysis and Retrieval of Text), the first computer based system was developed; only in the last couple years focus have been put into adapting the same techniques to mathematical content.

@@

\section{Mathematical based search projects}

>>>>

\section{Related Work}
\subsection{Status of Mathematical Information Retrieval}
Even though Information Retrieval has been a research field from around the 1950s where the first metrics for such systems were developed and in the 1960s where SMART (System for Mechanical Analysis and Retrieval of Text), the first computer based system was developed; only in the last couple years focus have been put into adapting the same techniques to mathematical content.

@@

\subsection{Mathematical based search projects}

### 2014-11-04 jcdl2015/paper/section-system_description.tex
First load of the JCDL paper files.

\chapter{$5e^{x+y}$(CDS)}
\label{chapter-cern_math_explorer}
In this chapter we present our solution called $5e^{x+y}$(CDS). We present the key points of our system, and how they are related to previous built systems.

\section{Features Extraction}
Our main goal is to develop a system that compliments the already available search capabilities in Invenio and in particular the CDS instance. Our system was developed on top of the  the Apache Lucene/Solr framework. CDS already uses Solr as a search engine layer, so developing $5e^{x+y}$(CDS) using this technology supposed a reasonable choice. One of the main steps while building an IR tool, is understanding the data and what kind of features can be extracted from it: If the system is based on text, then tasks such as stemming, handling of synonyms, misspelling and so on, should be taken into consideration; if the system being developed is based on images, then examples of features can be histograms of colors, geometrical features and so on. During the literature review, we discovered that most of the projects handled the notational aspects of the expressions, that is, the real naming of the terms in a given equation and each project developed its own set of heuristics to handle the structural component.

@@

\begin{longtable}{|m{4cm}|>{\centering\arraybackslash}m{2.15cm}|>{\centering\arraybackslash}m{1.65cm}|>{\centering\arraybackslash}m{1.65cm}|>{\centering\arraybackslash}m{1.65cm}|>{\centering\arraybackslash}m{1.65cm}|}
\hline 
\textbf{Character Name} & \textbf{Visual Rendering} & \textbf{NFC} & \textbf{NFD} & \textbf{NFKC} & \textbf{NFKD} \\ 
\hline 
LATIN CAPITAL LETTER A & A & 0x41 & 0x41 & 0x41 & 0x41 \\ 
\hline 
LATIN CAPITAL LETTER A WITH MACRON & \={A} & 0x100 & 0x41 0x304 & 0x100 & 0x41 0x304 \\
\hline 
LATIN CAPITAL LETTER A WITH RING ABOVE & \r{A} & 0xc5 & 0x41 0x30a & 0xc5 & 0x41 0x30a \\
\hline 
ANGSTROM SIGN (0x212b) &	\r{A} &	0xc5 &	0x41 0x30a &	0xc5 &	0x41 0x30a \\
\hline
LATIN CAPITAL LIGATURE IJ &	Ĳ &	0x132 &	0x132 &	0x49(I) 0x4a(J) &	0x49 0x4a \\
\hline
CYRILLIC CAPITAL LIGATURE A IE & {\unicodefont Ӕ} &	0x4d4 &	0x4d4 &	0x4d4 &	0x4d4 \\
\hline
LATIN CAPITAL LETTER AE	& {\unicodefont Ӕ}	& 0xc6	& 0xc6	& 0xc6	& 0xc6 \\
\hline
DOUBLE-STRUCK CAPITAL N	& {\unicodefont ℕ}	& 0x2115 & 0x2115 & 0x2115 & 0x4e \\
\hline
LATIN CAPITAL LETTER N	& N & 0x4e	& 0x4e & 	0x4e	& 0x4e \\
\hline 
MATHEMATICAL ITALIC CAPITAL N & $\mathcal{N}$	& 0xd835  0xdc41 &	0xd835  0xdc41 &	0x4e	 & 0x4e \\
\hline 
ROMAN NUMERAL TWELVE	& {\unicodefont Ⅻ}	&0x216b &	0x216b	& 0x58 0x49 0x49 &	0x58 0x49 0x49 \\
\hline
LATIN CAPITAL LETTER X - LATIN CAPITAL LETTER I - LATIN CAPITAL LETTER I &	XII &	0x58 0x49 0x49 &	0x58 0x49 0x49&	0x58(X) 0x49(I) 0x49(I) &	0x58 0x49 0x49 \\
\hline
VULGAR FRACTION ONE QUARTER	& \char"00BC &	0xbc &	0xbc &	0x31(1) 0x2044(/) 0x34(4) &	0x31 0x2044 0x34 \\
\hline
DIGIT ONE - SOLIDUS - DIGIT 4 &	1⁄4 &	0x31  0x2f 0x34 &	0x31 0x2f 0x34	& 0x31 0x2f 0x34 &	0x31 0x2f 0x34\\
\hline
DIGIT ONE - SOLIDUS - DIGIT 4 &	1⁄4 &	0x31  0x2f 0x34 &	0x31 0x2f 0x34	& 0x31 0x2f 0x34 &	0x31 0x2f 0x34\\
\hline
DIGIT ONE - FRACTION SLASH - DIGIT 4 & 1⁄4 &	0x31 0x2044 0x34 &	0x31 0x2044 0x34 &	0x31 0x2044 0x34 &	0x31 0x2044 0x34 \\
\hline
THERE EXISTS	& {\unicodefont ∃}&	0x2203 &	0x2203 &	0x2203 &	0x2203 \\
\hline
THERE DOES NOT EXIST	& {\unicodefont ∄} & 	0x2204	& 0x2203 0x338& 	0x2204 &	0x2203 0x338\\
\hline
LATIN CAPITAL LETTER OPEN E	& {\unicodefont Ɛ} &	0x190 &	0x190 &	0x190 &	0x190 \\
\hline
EULER CONSTANT	& {\unicodefont ℇ} & 0x2107 & 0x2107 & 0x190 & 0x190 \\
\hline
PLANCK CONSTANT	& {\unicodefont ℎ} &	0x210e &	0x210e &	0x68 &	0x68 \\
\hline
LATIN SMALL LETTER H &	h &	0x68 &	0x68 &	0x68 &	0x68 \\
\hline
PLANCK CONSTANT OVER TWO PI	& {\unicodefont ℏ} &	0x210f &	0x210f	& 0x127 &	0x127 \\
\hline
LATIN SMALL LETTER H WITH STROKE & ħ &	0x127	& 0x127	& 0x127	& 0x127 \\
\hline
GREEK CAPITAL LETTER OMEGA	& Ω &	0x3a9 &	0x3a9 &	0x3a9 &	0x3a9 \\
\hline
OHM SIGN & Ω &	0x2126 &	0x2126 &	0x3a9 &	0x3a9 \\
\hline
INTEGRAL & {\unicodefont ∫} &	0x222b &	0x222b &	0x222b &	0x222b \\
\hline
DOUBLE INTEGRAL	& {\unicodefont ∬} & 	0x222c &	0x222c &	0x222b 0x222b &	0x222b 0x222b \\
\hline
CONTOUR INTEGRAL	& {\unicodefont ∮} &	0x222e	 &0x222e &	0x222e &	0x222e \\
\hline
NABLA	& {\unicodefont ∇} &	0x2207	 &0x2207 &	0x2207 &	0x2207 \\
\hline
WHITE DOWN-POINTING TRIANGLE	& {\unicodefont ▽} &	0x25bd	 &0x25bd &	0x25bd &	0x25bd \\
\hline

\caption{Characters with similar glyphs and/or semantics}
\label{unicode_characters}
\end{longtable}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[height=2 cm]{figures/integrals.png}
	\caption{List of different types of integral related characters in Unicode}
	\label{integrals}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[height=2 cm]{figures/greater_less.png}
	\caption{List of different types of greater/less than related characters in Unicode}
	\label{less_greater}
\end{figure}

@@

\section{Integration with Mathematica}
This section describes some aspects that had to be considered while implementing the integration with Mathematica. Our system uses the {\codefont JLink} library to create a connection with the underlying process {\codefont MathKernel}. The main class for communicating java code and {\codefont Mathematica} is {\codefont KernelLink} and the main used method was {\codefont evaluateToOutputForm(String expr, int pageWidth)} where {\codefont expr} is the expression to be evaluated and {\codefont pageWidth} affects the formatting of the result.
Once all the boilerplate has been written, the steps followed to define and find patterns in the expression are as follow: We import the string in MathML format through the command {\codefont ImportString["<mathmlExpression>", "MathML"]}. This creates the expression in the box form, but no further analysis is possible from this form. We then try to interpret this form into a fully Mathematica compatible expression by using {\codefont MakeExpression}

@@

\begin{longtable}{|>{\centering\arraybackslash}m{4.5cm}|>
{\centering\arraybackslash}m{4.5cm}|>
{\centering\arraybackslash}m{4.5cm}|}
\hline
\small{Original Form} & \small{{\codefont Simplify}} & \small{{\codefont FullSimplify}} \\
\hline
\scriptsize{\codefont <math> <mrow> <mrow> <mfrac> <msup> <mi>Φ</mi> <mo>′</mo> </msup> <mrow> <mn>2</mn> <mo>⁢</mo> <mrow> <mo>(</mo> <mrow> <mn>1</mn> <mo>+</mo> <mi>Φ</mi> </mrow> <mo>)</mo> </mrow> </mrow> </mfrac> <mo>+</mo> <mfrac> <mn>1</mn> <mi>r</mi> </mfrac> </mrow> <mo>=</mo> <mn>0</mn> </mrow> </math> }      & \scriptsize{\codefont <math> <mrow> <mrow> <mfrac> <mn>1</mn> <mi>r</mi> </mfrac> <mo>+</mo> <mfrac> <msup> <mi>Φ</mi> <mo>′</mo> </msup> <mrow> <mrow> <mn>2</mn> <mo>⁢</mo> <mi>Φ</mi> </mrow> <mo>+</mo> <mn>2</mn> </mrow> </mfrac> </mrow> <mo></mo> <mn>0</mn> </mrow> </math>}    & \scriptsize{\codefont <math> <mrow> <mrow> <mfrac> <mn>2</mn> <mi>r</mi> </mfrac> <mo>+</mo> <mfrac> <msup> <mi>Φ</mi> <mo>′</mo> </msup> <mrow> <mi>Φ</mi> <mo>+</mo> <mn>1</mn> </mrow> </mfrac> </mrow> <mo></mo> <mn>0</mn> </mrow> </math>}     \\
$\frac{\Phi '}{2 (1+\Phi )}+\frac{1}{r}=0$          & $\frac{1}{r}+\frac{\Phi '}{2 \Phi +2}=0$ & $\frac{2}{r}+\frac{\Phi '}{\Phi +1}=0$       \\

\hline

\scriptsize{\codefont <math> <mrow> <mrow> <mfrac> <mn>1</mn> <mn>2</mn> </mfrac> <mo>⁢</mo> <msup> <mi>x</mi> <mn>2</mn> </msup> <mo>⁢</mo> <mrow> <mi>log</mi> <mo>⁡</mo> <mrow> <mo>(</mo> <mrow> <mn>1</mn> <mo>-</mo> <msup> <mi>e</mi> <mrow> <mo>-</mo> <mrow> <mn>2</mn> <mo>⁢</mo> <mi>x</mi> </mrow> </mrow> </msup> </mrow> <mo>)</mo> </mrow> </mrow> </mrow> <mo>-</mo> <mrow> <mfrac> <mn>1</mn> <mn>2</mn> </mfrac> <mo>⁢</mo> <mi>x</mi> <mo>⁢</mo> <mi>L</mi> <mo>⁢</mo> <msub> <mi>i</mi> <mn>2</mn> </msub> <mo>⁢</mo> <mrow> <mo>(</mo> <msup> <mi>e</mi> <mrow> <mo>-</mo> <mrow> <mn>2</mn> <mo>⁢</mo> <mi>x</mi> </mrow> </mrow> </msup> <mo>)</mo> </mrow> </mrow> <mo>-</mo> <mrow> <mfrac> <mn>1</mn> <mn>4</mn> </mfrac> <mo>⁢</mo> <mi>L</mi> <mo>⁢</mo> <msub> <mi>i</mi> <mn>3</mn> </msub> <mo>⁢</mo> <mrow> <mo>(</mo> <msup> <mi>e</mi> <mrow> <mo>-</mo> <mrow> <mn>2</mn> <mo>⁢</mo> <mi>x</mi> </mrow> </mrow> </msup> <mo>)</mo> </mrow> </mrow> </mrow> </math> }      & \scriptsize{\codefont <math> <mrow> <mfrac> <mn>1</mn> <mn>4</mn> </mfrac> <mo>⁢</mo> <msup> <mi>e</mi> <mrow> <mrow> <mo>-</mo> <mn>2</mn> </mrow> <mo>⁢</mo> <mi>x</mi> </mrow> </msup> <mo>⁢</mo> <mrow> <mo>(</mo> <mrow> <mrow> <mn>2</mn> <mo>⁢</mo> <msup> <mi>x</mi> <mn>2</mn> </msup> <mo>⁢</mo> <msup> <mi>e</mi> <mrow> <mn>2</mn> <mo>⁢</mo> <mi>x</mi> </mrow> </msup> <mo>⁢</mo> <mrow> <mi>log</mi> <mo>⁡</mo> <mo>(</mo> <mrow> <mn>1</mn> <mo>-</mo> <msup> <mi>e</mi> <mrow> <mrow> <mo>-</mo> <mn>2</mn> </mrow> <mo>⁢</mo> <mi>x</mi> </mrow> </msup> </mrow> <mo>)</mo> </mrow> </mrow> <mo>-</mo> <mrow> <mn>2</mn> <mo>⁢</mo> <msub> <mi>i</mi> <mn>2</mn> </msub> <mo>⁢</mo> <mi>L</mi> <mo>⁢</mo> <mi>x</mi> </mrow> <mo>-</mo> <mrow> <msub> <mi>i</mi> <mn>3</mn> </msub> <mo>⁢</mo> <mi>L</mi> </mrow> </mrow> <mo>)</mo> </mrow> </mrow> </math> }    & \scriptsize{\codefont <math> <mrow> <mrow> <mfrac> <mn>1</mn> <mn>2</mn> </mfrac> <mo>⁢</mo> <msup> <mi>x</mi> <mn>2</mn> </msup> <mo>⁢</mo> <mrow> <mi>log</mi> <mo>⁡</mo> <mo>(</mo> <mrow> <mn>1</mn> <mo>-</mo> <msup> <mi>e</mi> <mrow> <mrow> <mo>-</mo> <mn>2</mn> </mrow> <mo>⁢</mo> <mi>x</mi> </mrow> </msup> </mrow> <mo>)</mo> </mrow> </mrow> <mo>-</mo> <mrow> <mfrac> <mn>1</mn> <mn>4</mn> </mfrac> <mo>⁢</mo> <mi>L</mi> <mo>⁢</mo> <msup> <mi>e</mi> <mrow> <mrow> <mo>-</mo> <mn>2</mn> </mrow> <mo>⁢</mo> <mi>x</mi> </mrow> </msup> <mo>⁢</mo> <mrow> <mo>(</mo> <mrow> <mrow> <mn>2</mn> <mo>⁢</mo> <msub> <mi>i</mi> <mn>2</mn> </msub> <mo>⁢</mo> <mi>x</mi> </mrow> <mo>+</mo> <msub> <mi>i</mi> <mn>3</mn> </msub> </mrow> <mo>)</mo> </mrow> </mrow> </mrow> </math>}     \\ 
\scriptsize{ $\frac{1}{2} x^2 \log \left(1-e^{-2 x}\right) - \frac{1}{2} x L i_ 2 \left(e^{-2 x}\right)-\frac{1}{4} L i_ 3 \left(e^{-2 x}\right)$ }          & \scriptsize { \begin{multline}  \frac{1}{4} e^{-2 x} (2 x^2 e^{2 x} \log (1-e^{-2 x})- \notag\\ 2 i_ 2 L x-i_ 3 L)  \end{multline} } & \scriptsize{ $\frac{1}{2} x^2 \log \left(1-e^{-2 x}\right)-\frac{1}{4} L e^{-2 x} \left(2 i_ 2 x+i_ 3\right)$ }      \\
\hline
\caption{Comparison of different simplification modes}

\label{comparison_simplification}

\end{longtable}
\subsection{Recursive interpretation}
Unfortunately, Mathematica cannot always interpret correctly an input expression. In this case, it can, however interpret some sub-expressions from the original one. Our system takes advantage of this, and the first attempt is to try to interpret the original string. If this step is not successful (the result variable equals the string "\{\}"), we identify the occurrences of {\codefont RowBox} which identify all the subexpressions in the tree. For each of this elements we call the {\codefont MakeExpression} and if it was successful we search for structural patterns in it. A first thought would tell us that if sub-tree was interpreted successfully, there is no need to try to go further down in that sub-tree, however, even if a sub-expression is interpreted correctly, the result sometimes is not complete and processing further down gives better results. For this reason, we do not discard any of the subexpressions and try to interpret all of them. For this same reason, we do not store the number of occurrences of a given pattern, but only whether it contains is or not.

@@

\begin{figure}[h!]
\includegraphics[height=4.2 cm]{visio_drawings/architecture.png}
\label{system_arch}
\caption{High level deployment architecture}
\end{figure}

@@

\begin{figure}
\center
\includegraphics[height=6.2 cm]{figures/input_interface.png}
\label{input_int}
\caption{Web interface of $5e^{x+y}$: Query input screen}
\end{figure}

\begin{figure}
\center
\includegraphics[height=6.2 cm]{figures/results_interface.png}
\label{results_int}
\caption{Web interface of $5e^{x+y}$: Results screen}
\end{figure}

>>>>

\section{$5e^{x+y}$(CDS)}
\label{chapter-cern_math_explorer}
In this chapter we present our solution called $5e^{x+y}$(CDS). We present the key points of our system, and how they are related to previous built systems.

\subsection{Features Extraction}
Our main goal is to develop a system that compliments the already available search capabilities in Invenio and in particular the CDS instance. Our system was developed on top of the  the Apache Lucene/Solr framework. CDS already uses Solr as a search engine layer, so developing $5e^{x+y}$(CDS) using this technology supposed a reasonable choice. One of the main steps while building an IR tool, is understanding the data and what kind of features can be extracted from it: If the system is based on text, then tasks such as stemming, handling of synonyms, misspelling and so on, should be taken into consideration; if the system being developed is based on images, then examples of features can be histograms of colors, geometrical features and so on. During the literature review, we discovered that most of the projects handled the notational aspects of the expressions, that is, the real naming of the terms in a given equation and each project developed its own set of heuristics to handle the structural component.

@@

@@

\begin{figure}[h!]
	\centering
	\includegraphics[height=2 cm]{images/figures/integrals.png}
	\caption{List of different types of integral related characters in Unicode}
	\label{integrals}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[height=2 cm]{images/figures/greater_less.png}
	\caption{List of different types of greater/less than related characters in Unicode}
	\label{less_greater}
\end{figure}

@@

\subsection{Integration with Mathematica}
This section describes some aspects that had to be considered while implementing the integration with Mathematica. Our system uses the {\codefont JLink} library to create a connection with the underlying process {\codefont MathKernel}. The main class for communicating java code and {\codefont Mathematica} is {\codefont KernelLink} and the main used method was {\codefont evaluateToOutputForm(String expr, int pageWidth)} where {\codefont expr} is the expression to be evaluated and {\codefont pageWidth} affects the formatting of the result.
Once all the boilerplate has been written, the steps followed to define and find patterns in the expression are as follow: We import the string in MathML format through the command {\codefont ImportString["<mathmlExpression>", "MathML"]}. This creates the expression in the box form, but no further analysis is possible from this form. We then try to interpret this form into a fully Mathematica compatible expression by using {\codefont MakeExpression}

@@

\subsection{Recursive interpretation}
Unfortunately, Mathematica cannot always interpret correctly an input expression. In this case, it can, however interpret some sub-expressions from the original one. Our system takes advantage of this, and the first attempt is to try to interpret the original string. If this step is not successful (the result variable equals the string "\{\}"), we identify the occurrences of {\codefont RowBox} which identify all the subexpressions in the tree. For each of this elements we call the {\codefont MakeExpression} and if it was successful we search for structural patterns in it. A first thought would tell us that if sub-tree was interpreted successfully, there is no need to try to go further down in that sub-tree, however, even if a sub-expression is interpreted correctly, the result sometimes is not complete and processing further down gives better results. For this reason, we do not discard any of the subexpressions and try to interpret all of them. For this same reason, we do not store the number of occurrences of a given pattern, but only whether it contains is or not.

@@

\begin{figure}[h!]
\includegraphics[height=3 cm]{images/visio_drawings/architecture.png}
\label{system_arch}
\caption{High level deployment architecture}
\end{figure}

@@

\begin{figure}
\center
\includegraphics[height=3 cm]{images/figures/input_interface.png}
\label{input_int}
\caption{Web interface of $5e^{x+y}$: Query input screen}
\end{figure}

\begin{figure}
\center
\includegraphics[height=3 cm]{images/figures/results_interface.png}
\label{results_int}
\caption{Web interface of $5e^{x+y}$: Results screen}
\end{figure}
